Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-26 TACAS2018/paper.tex
ch 4

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big\{ \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%
Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}) \, \forall k\in [0,n-1], \,
\vec{x}_0 \in X_0, \,
\vec{u}_{k} \in U  \} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and vector sums preserve convexity \note{ref}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$ 
generalises the previous notions over an unbounded time horizon (transitive closure).

@@

In many models of real-world control systems, 
the internal state $x$ cannot be directly measured,  
but is rather inferred from its effect on the output $\vec{y}$ using an observer.  
Essentially, a state observer is another dynamical model (denoted below with variable $\hat{\vec{x}}$) 
that provides an estimate of the internal state $x$ of the given model, 
from measurements of the input and output of the system (See~\ref{sec:observable}).

A closed-loop digital control system with observer is shown in Figure~\ref{fig:observersystem}. 
The feedback controller is driven by the observer state $\hat{\vec{x}}$, 
and is characterised by a gain matrix $\mat{K}_{\langle C \rangle}$ that influences the closed-loop dynamics.  
Moreover, $\vec{y}_e=\hat{\vec{y}}_k - \vec{y}_k$ is the output estimation error,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
  whereas $\mat{L}_{\langle C \rangle}$ is an error gain matrix that amplifies the output estimation error. 
Figure~\ref{fig:observersystem} also contains references to $\nu_1$, $\nu_2$, $\nu_1'$, $\nu_2'$ and $\nu_t$, which
  denote noise terms that will be explained later.   
%%   With respect to noises (denoted by $\nu$), 
%% $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations.
  Lastly, $k_r$ is a reference gain required by the model.  
  Our goal in this work is to synthesise both $\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$.

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++ 
 The time discretization in Sec.~\ref{sec:model} only 
 allows us to evaluate the state of the system at integer multiples of the sample time.
 However, in order to synthesise $\mat{K}=\mat{K}_{\langle C \rangle}$ and $\mat{L}=\mat{L}_{\langle C \rangle}$, 
 we must be able to ensure that certain properties of the plant hold at any given real time.
 In order to do so, 
 we accelerate the dynamics of the system according to the model semantics in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$  at a given time $T$, 
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time $t=T$.

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for continuous-time models without inputs}\label{sec:cont_acc_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

 \begin{theorem}
 The solution to the differential equation $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$,
 evaluated at time $T$ is
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x} (0)\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(2)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&\quad gm(\lambda_i)>1\\0&\quad gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  
%$\vec{x}_T$ is the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x} \text{ with } T \in \mathbb{R}^+$ is the state of the system at time $t=T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in studying models with parametric inputs, 
 which are described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where $\forall t\geq 0, \mat{B}\vec{u}(t)=\vec{u}(0)$. 
 We will assume $\mat{A}$ is invertible when there are geometric multiplicities $>1$.
 For such models, a the state $\vec{x}_T$ is computed next.  
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} \text{ where } \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn}, 
 is the state of the model $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}\vec{u}) \text{ with } T \in \mathbb{R}^+$ is the state of the system at time $T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big\{ \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%
Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and vector sums preserve convexity \note{ref}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$ 
generalises the previous notions over an unbounded time horizon (transitive closure).

@@

In many models of real-world control systems, 
the internal state $x$ cannot be directly measured,  
but is rather inferred from its effect on the output $\vec{y}$ using an observer~~\cite{Astrom08}.  
Essentially, a state observer is a new dynamical model (with variable denoted below as $\hat{\vec{x}}$) 
that provides an estimate of the internal state $x$ of the given model, 
from measurements of its inputs and outputs (See~\ref{sec:observable}).

A closed-loop digital control system with observer is shown in Figure~\ref{fig:observersystem}. 
The linear feedback controller *bottom left) and is characterised by a gain matrix $\mat{K}_{\langle C \rangle}$ that influences the closed-loop dynamics, 
and is driven by the observer state $\hat{\vec{x}}$.  
Moreover, $\vec{y}_e=\hat{\vec{y}} - \vec{y}$ is the output estimation error,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
 which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$. 
Figure~\ref{fig:observersystem} also contains references to $\nu_1$, $\nu_2$, $\nu_1'$, $\nu_2'$ and $\nu_t$, which
  denote noise terms that will be explained later.   
%%   With respect to noises (denoted by $\nu$), 
%% $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations.
  Lastly, $k_r$ is a reference gain required by the model.  
  Our goal in this work is to synthesise both $\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$, 
  so that specific closed-loop properties of the overall interconnection are satisfied.

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++ 
 The time discretization in Sec.~\ref{sec:model} only 
 allows us to evaluate the state of the plant at integer multiples of the sample time.
 However, in order to synthesise $\mat{K}=\mat{K}_{\langle C \rangle}$ and $\mat{L}=\mat{L}_{\langle C \rangle}$, 
 we must be able to ensure that certain properties of the plant hold at any given real time.
 In order to do so, 
 we accelerate the dynamics of the plant according to the model semantics in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$  at a given time $T$, 
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time $t=T$.

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated\! dynamics\! for\! continuous-time\! models\! without\! inputs}\label{sec:cont_acc_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

 \begin{theorem}
 The solution to the differential equation $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$,
 evaluated at time $T$ is
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x} (0)\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(2)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&\quad gm(\lambda_i)>1\\0&\quad gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  
%$\vec{x}_T$ is the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x} \text{ with } T \in \mathbb{R}^+$ is the state of the system at time $t=T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in studying models with parametric inputs, 
 which are described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where $\forall t\geq 0, \mat{B}\vec{u}(t)=\vec{u}$. 
 We will assume $\mat{A}$ is invertible when there are non-trivial geometric multiplicities.
 For such models, a the state $\vec{x}_T$ is computed next.  
\begin{theorem}
The state of the model $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$ is 
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} 
 \end{align}
 \note{can you simplify in a single line?}
 where $\forall t \leq T,\ \vec{u}(t)=\vec{u}$,  
 and $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn}.  
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists \note{there always exist one, since you define it from A} a discrete matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, 
 then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}\vec{u}) \text{ with } T \in \mathbb{R}^+$ \note{the formula appears to not be working for T=0 - why? Also, it is not clear what the real-power of a matrix A is?} is the state of the system at time $T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
